\mytitle

### 2015-07-27 MSc_diss.tex
Continued subsection about model validation

@@

\subsection{Chemical mechanism}
The chemical mechanism of the model is based on recommendations from \citep{Atkinson2004} and information from the Master Chemical Mechanism, MCM v3.3 \citep{Jenkin1997,Saunders2003}, retrieved via website: http://mcm.leeds.ac.uk/MCM. On a whole it includes 90 species and 165 reactions. The mechanism is developed to represent the chemistry that governs the production and loss of ozone, hydrocarbons and alkyl nitrates. These include:

@@

@@

Two experiments have been performed. The aim of the first one is to reveal the total impact of alkyl nitrate chemistry on $O_3$ production, the aim of the second one is to estimate the contribution of individual alkyl nitrates to changes in $O_3$ production.

@@

For the second experiment the model was run using the same approach concerning variations in $NO_x$ and $VOCs$ concentrations, but changes in alkyl nitrate chemistry were treated differently. In every model run reactions of individual alkyl nitrates or their groups (defined in Table\ref{tab:RHandANS} were either switched on or switched off, resulting in other additional 847 model runs.

@@

\begin{table} % model setup: O3, CO
\caption{Fixed and variable parameters used in box model runs}\label{tab:setupO3CO}
\centering
\begin{tabular}{ccc}
\hline
Parameter & Initial mixing ratio & Type \\
\hline
$H_2O$    & 2\%                  & Fixed    \\
$O_2$     & 21\%                 & Fixed    \\
$N_2$     & 78\%                 & Fixed    \\
$H_2$     & 525 ppb              & Variable \\
$O_3$     & 40 ppb               & Variable \\
$CO$      & 100 ppb              & Variable \\
\hline
\end{tabular}
\end{table}

@@

\begin{table} % model setup: VOCs
\caption{Model setup: VOCs concentrations (ppb)}\label{tab:setupVOCs}
\centering
\begin{tabular}{ccccccccc}
\hline
Experiment number & $CH_4$ & $C_2H_6$ & $C_3H_8$ & $nC_4H_{10}$ & $iC_4H_{10}$ & $nC_5H_{12}$ & $iC_5H_{12}$ \\
\hline
1  & 1780 &	0.620 &	0.100 &	0.050 &	0.250 &	0.250 &	0.250 \\
2  & 1807 &	1.258 &	0.390 &	0.245 &	0.325 &	0.275 &	0.425 \\
3  & 1834 &	1.896 &	0.680 &	0.440 &	0.400 &	0.300 &	0.600 \\
4  & 1861 &	2.534 &	0.970 &	0.635 &	0.475 &	0.325 &	0.775 \\
5  & 1888 &	3.172 &	1.260 &	0.830 &	0.550 &	0.350 &	0.950 \\
6  & 1915 &	3.810 &	1.550 &	1.025 &	0.625 &	0.375 &	1.125 \\
7  & 1942 &	4.448 &	1.840 &	1.220 &	0.700 &	0.400 &	1.300 \\
8  & 1969 &	5.086 &	2.130 &	1.415 &	0.775 &	0.425 &	1.475 \\
9  & 1996 &	5.724 &	2.420 &	1.610 &	0.850 &	0.450 &	1.650 \\
10 & 2023 &	6.362 &	2.710 &	1.805 &	0.925 &	0.475 &	1.825 \\
11 & 2050 &	7.000 &	3.000 &	2.000 &	1.000 &	0.500 &	2.000 \\
\hline
\end{tabular}
\end{table}

\begin{figure}[h]
\includegraphics[width=0.5\linewidth]{{./pics/Sillman1999_Fig1}.png}
%\caption{chem_allAN_netO3_rate}
\label{fig:O3observations}
\end{figure}

\section{Results} \label{sec:res}
Dependence of alkyl nitrate yield on carbon number (paper about n-alkanes): bigger yield of ANs from larger carbon number means that the potential for contributing to photochemical air pollution may be less for the larger (C>6) n-alkanes that for the smaller ones.

\subsection{Justification of model performance}
!! from \citep{Finlayson-Pitts2000}
Urban:
Certain reproducible features of time-concentration profiles for  pollutants are observed in "smoggy" ambient air. Figure 1.3, a classic example of historical interest, shows such profiles for NO, NO2, and total oxidant (mainly O3) in Pasadena, California, during a  severe photochemical air pollution episode in July 1973.
Reproducible  features  include  the  following:
\begin{itemize}
\item In the early morning, the concentration of NO rises and  reaches a maximum at a time that approximately coincides with the  maximum emissions of NO, in this case, peak  automobile  traffic; 
\item Subsequently, NO2 rises to a maximum;
\item Oxidant (e.g., O3) levels, which are relatively low in the  early morning, increase significantly about noon when the NO concentration drops to a low value.
\end{itemize}

\begin{figure}[h]
\includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3_rate.png}
%\caption{chem_allAN_netO3_rate}
\label{fig:O3observations}
\end{figure}

@@

\subsection{Short-chained alkyl nitrates} \label{sec:short_an}
\subsubsection{Short photochemical processing times} \label{sec:short_an_short_time}

\section{Conclusion} \label{sec:conclusion}

\newpage
\clearpage
\thispagestyle{empty}% empty page style (?)
\begin{landscape}% Landscape page
\centering
	%\input{subfig_28_land}
\end{landscape}

\bibliography{diss_refs}

>>>>

! \citep{Newland2013}
[The model is written using the programming language FACSIMILE (Curtis and Sweetenham, 1987) and the differential equations are assembled and integrated using the Gear’s method (1971). The mixing ratios of the input species in the model are driven entirely by the emissions, the loss processes and the transport with no redistribution of the burden to ensure that mixing ratios agree with observations.]

@@

\subsection{Chemical mechanism}
The chemical mechanism of the model is based on recommendations from \citep{Atkinson2004} and information from the Master Chemical Mechanism, MCM v3.3 \citep{Jenkin1997,Saunders2003}, retrieved via website: http://mcm.leeds.ac.uk/MCM. On a whole it includes 90 species and 165 reactions. The mechanism is developed to represent the chemistry that governs the production and loss of ozone, hydrocarbons and alkyl nitrates. These include:

! \citep{Newland2013}
[The majority of the kinetic reaction rates and photochemical data used in the full chemistry model were taken from the IUPAC Subcommittee for Gas Kinetic Data Evaluation (http://www.iupac-kinetic.ch.cam.ac.uk; Atkinson et al., 2004; Atkinson et al., 2006; Atkinson et al., 2008). If the required data were not provided by IUPAC then data were taken from the NASA / JPL ‘Chemical Kinetics and Photochemical Data for Use in Atmospheric Studies: Evaluation No. 17’ (Sander et al., 2011). Both these sources consider the full range of literature available on each reaction and provide a recommendation based on a synthesis of this data. There are sometimes minor differences between the recommendations of each organisation.]

@@

\citep{Newland2013}
[The alkanes and alkyl nitrates studied in this work were all explicitly described in the model with oxidation of the alkane giving a peroxy radical and reaction of the peroxy radical with NO giving an alkyl nitrate.]

@@

Two experiments have been performed. The aim of the first one is to reveal the total impact of alkyl nitrate chemistry on $O_3$ production, the aim of the second one is to estimate the contribution of individual alkyl nitrates to changes in $O_3$ production (sensitivity!).

@@

For the second experiment the model was run using the same approach concerning variations in $NO_x$ and $VOCs$ concentrations, but changes in alkyl nitrate chemistry were treated differently. In every model run reactions of individual alkyl nitrates or their groups (defined in Table \ref{tab:RHandANS}) were either switched on or switched off, resulting in other 847 additional model runs.

@@

\begin{table} % model setup: O3, CO
\caption{Fixed and variable 'independent' parameters used in box model runs}\label{tab:setupO3CO}
\centering
\begin{tabular}{ccc}
\hline
Parameter & Initial mixing ratio & Type \\
\hline
$H_2O$    & 2\%                  & Fixed    \\
$O_2$     & 21\%                 & Fixed    \\
$N_2$     & 78\%                 & Fixed    \\
$H_2$     & 525 ppb              & Variable \\
$O_3$     & 40 ppb               & Variable \\
$CO$      & 100 ppb              & Variable \\
\hline
\end{tabular}
\end{table}

@@

\begin{table} % model setup: VOCs
\caption{Model setup: VOCs concentrations (ppb)}\label{tab:setupVOCs}
\centering
\begin{tabular}{ccccccccc}
\hline
Experiment number & $CH_4$ & $C_2H_6$ & $C_3H_8$ & $nC_4H_{10}$ & $iC_4H_{10}$ & $nC_5H_{12}$ & $iC_5H_{12}$ \\
\hline
A  & 1780 &	0.620 &	0.100 &	0.050 &	0.250 &	0.250 &	0.250 \\
B  & 1807 &	1.258 &	0.390 &	0.245 &	0.325 &	0.275 &	0.425 \\
C  & 1834 &	1.896 &	0.680 &	0.440 &	0.400 &	0.300 &	0.600 \\
D  & 1861 &	2.534 &	0.970 &	0.635 &	0.475 &	0.325 &	0.775 \\
E  & 1888 &	3.172 &	1.260 &	0.830 &	0.550 &	0.350 &	0.950 \\
F  & 1915 &	3.810 &	1.550 &	1.025 &	0.625 &	0.375 &	1.125 \\
G  & 1942 &	4.448 &	1.840 &	1.220 &	0.700 &	0.400 &	1.300 \\
H  & 1969 &	5.086 &	2.130 &	1.415 &	0.775 &	0.425 &	1.475 \\
I  & 1996 &	5.724 &	2.420 &	1.610 &	0.850 &	0.450 &	1.650 \\
J  & 2023 &	6.362 &	2.710 &	1.805 &	0.925 &	0.475 &	1.825 \\
K  & 2050 &	7.000 &	3.000 &	2.000 &	1.000 &	0.500 &	2.000 \\
\hline
\end{tabular}
\end{table}
\subsection{Model validation}
Time evolution of species concentrations in the model agrees reasonably well with photochemical theory as it can be seen, for example, from Figure ???. At low $NO_x$ levels (50 ppt) $O_3$ is being removed from the system because there is not enough $NO$ that could react with the peroxy radicals ($HO_2$ and $RO_2$) to produce $NO_2$ (which photolysis in conjunction with the reaction $O(^3P)+O_2\xrightarrow{M} O_3$ is the main source of $O_3$ in the troposphere). Some additional $O_3$ removal occurs because $HO_2$ radicals can react with $O_3$ leading to the regeneration of $OH$ radicals as part of an $O_3$ depleting $OH$-$HO_2$ ($HO_x$) cycle. In other words, the following reaction sequence dominates:
\begin{itemize}
\item[] $O_3 + hv \rightarrow O(^1D) + O_2$
\item[] $O(^1D) + H_2O \rightarrow OH + OH$
\item[] $O(^1D) \xrightarrow{M} O(^3P)$
\item[] $O(^3P) + O_2 \xrightarrow{M} O_3$
\item[] $OH + CO \rightarrow HO_2 + CO_2$
\item[] $OH + RH \rightarrow RO_2 + H_2O$
\item[] $HO_2 + HO_2 \rightarrow H_2O_2 + O_2$
\item[] $RO_2 + HO_2 \rightarrow ROOH + O_2$
\item[] $O_3 + HO_2 \rightarrow OH + 2O_2$.
\end{itemize}
A more pronounced $O_3$ loss takes place under lower $VOC$ concentrations because in this case more $OH$ radicals are left in the system after oxidation of $CO$ and alkanes leading to a greater concentration of $HO_2$ radicals, and therefore to a greater chance of the last reaction in the sequence.

The concentration of NOx is the main factor which determines whether O3 is produced or removed in the atmosphere.
changing balance of the free-radical reactions, and their impact on O3, as the level of NOx increases.

NO and NO2 exist in a state equilibrium with rapid interconversion between the two molecules in the troposphere.

!! from \citep{Finlayson-Pitts2000}
Urban:
Certain reproducible features of time-concentration profiles for  pollutants are observed in "smoggy" ambient air. Figure 1.3, a classic example of historical interest, shows such profiles for NO, NO2, and total oxidant (mainly O3) in Pasadena, California, during a  severe photochemical air pollution episode in July 1973.
Reproducible  features  include  the  following:
\begin{itemize}
\item In the early morning, the concentration of NO rises and  reaches a maximum at a time that approximately coincides with the  maximum emissions of NO, in this case, peak  automobile  traffic; 
\item Subsequently, NO2 rises to a maximum;
\item Oxidant (e.g., O3) levels, which are relatively low in the  early morning, increase significantly about noon when the NO concentration drops to a low value.
\end{itemize}

%\newpage
%\clearpage
%\thispagestyle{empty}% empty page style (?)
%\begin{landscape}% Landscape page
%\centering
%	\input{diss_valid_vs_time}
%\end{landscape}

\section{Results} \label{sec:res}
\begin{figure}[h] % Sillman1999
\includegraphics[width=0.5\linewidth]{{./pics/Sillman1999_Fig1_adapted}.png}
\caption{Isopleths giving net rate of ozone production (ppb/h, solid lines) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) for mean summer daytime meteorology and clear skies. The solid lines represent production rates of 1, 2.5, 5, 10, 15, 20 and 30 ppb/h. Adapted from \citep{Sillman1999}.}\label{fig:Sillman1999}
\end{figure}
\begin{figure}[h] % allAN_netO3_rate
\includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3_rate.png}
\caption{allAN, netO3 rate}\label{fig:allAN_netO3_rate}
\end{figure}

! Dependence of alkyl nitrate yield on carbon number (paper about n-alkanes): bigger yield of ANs from larger carbon number means that the potential for contributing to photochemical air pollution may be less for the larger (C>6) n-alkanes that for the smaller ones.

@@

\section{Conclusion} \label{sec:conclusion}
\bibliography{diss_refs}
